Catastrophic Goodhart: regularizing RLHF with KL divergence does not mitigate heavy-tailed reward misspecification

### 2024-10-24 paper/supplementary/theorem5.tex
add icml-mi version of paper

\documentclass[../neurips_2024.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}

@@

 The conditional expectation \(\mathbb E[V | X + V > t]\) is given by \(\frac{\int_{-\infty}^\infty vf_V(v)\text{Pr}(X>t-v)} {\int_{-\infty}^\infty f_V(v)\text{Pr}(X>t-v)}\), \footnote{We'll generally omit \(dx\) and \(dv\) terms in the interests of compactness and laziness; the implied differentials should be pretty clear.}
 and we divide the integral in the numerator into 4 regions, showing that each region's effect on the conditional expectation of V is similar to that of the corresponding region in the unconditional expectation \(\mathbb E[V]\).

The regions are defined in terms of a slow-growing function \(h(t):\mathbb R \to \mathbb R_{\ge 0}\) such that the fiddly bounds on different pieces of the proof work out. Roughly, we want it to go to infinity so that \(|V|\) is likely to be less than \(h(t)\) in the limit, but grow slowly enough that the shape of \(V\) 's distribution within the interval \([-h(t),h(t)]\) doesn't change much after conditioning.

@@

Formally, this means that \(\lim_{x\to\infty}\frac{\text{Pr}(X_1+X_2>x)}{\text{Pr}(X>x)} = 2\). 
This happens roughly whenever \(X\) has tails that are heavier than \(e^{-cx}\) for any \(c\) and is reasonably well-behaved; counterexamples to the claim "long-tailed implies subexponential" exist, but they're nontrivial to exhibit.
Examples of subexponential distributions include log-normal distributions, anything that decays like a power law, the Pareto distribution,and distributions with tails asymptotic to \(e^{-x^a}\) for any \(0<a<1\).

@@

With all that out of the way, we can move on to the proof.

@@

Note that \(\int_{t}^\infty vf_V(v) = t\int_{t}^\infty f_V(v) + \int_{t}^\infty (v-t)f_V(v) = t\bar F_V(t)+\int_{t}^\infty \bar F_V(v)\), so it will suffice to show that both terms in this sumare \(o(\bar F_X(t))\).

>>>>

\documentclass[../main.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}

@@

 The conditional expectation \(\mathbb E[V | X + V > t]\) is given by \(\frac{\int_{-\infty}^\infty vf_V(v)\text{Pr}(X>t-v)} {\int_{-\infty}^\infty f_V(v)\text{Pr}(X>t-v)}\), \footnote{We'll generally omit \(dx\) and \(dv\) terms in the interests of compactness and conciseness; the implied differentials should be pretty clear.}
 and we divide the integral in the numerator into 4 regions, showing that each region's effect on the conditional expectation of V is similar to that of the corresponding region in the unconditional expectation \(\mathbb E[V]\).

The regions are defined in terms of a slow-growing function \(h(t):\mathbb R \to \mathbb R_{\ge 0}\) such that the fiddly bounds on different pieces of the proof work out. Roughly, we want it to go to infinity so that \(|V|\) is likely to be less than \(h(t)\) in the limit, but grow slowly enough that the shape of \(V\)'s distribution within the interval \([-h(t),h(t)]\) doesn't change much after conditioning.

@@

Formally, this means that \(\lim_{x\to\infty}\frac{\text{Pr}(X_1+X_2>x)}{\text{Pr}(X>x)} = 2\). 
This occurs roughly whenever \(X\) has tails that are heavier than \(e^{-cx}\) for any \(c\) and is reasonably well-behaved; counterexamples to the claim "long-tailed implies subexponential" exist, but they're nontrivial to exhibit.
Examples of subexponential distributions include log-normal distributions, anything that decays like a power law, the Pareto distribution,and distributions with tails asymptotic to \(e^{-x^a}\) for any \(0<a<1\).

@@

With these definitions and assumptions, we can move on to the proof.

@@

Note that \(\int_{t}^\infty vf_V(v) = t\int_{t}^\infty f_V(v) + \int_{t}^\infty (v-t)f_V(v) = t\bar F_V(t)+\int_{t}^\infty \bar F_V(v)\), so it will suffice to show that both terms in this sum are \(o(\bar F_X(t))\).
